Estimation of Intracortical Connectivity in a Dynamical Neural Field Model

### 2010-06-24 ltx/PLoS/LaTeX_Manuscript.tex
edits to figure captions

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{A Data Driven Framework for Patient-Specific Neural Field Modelling}
}
% Insert Author names, affiliations and corresponding author email.
\\
Dean R. Freestone$^{1,2,3}$, 
Parham Aram$^{4}$, 
Michael Dewar$^{5,\ast}$,
Kenneth Scerri$^{6,\ast}$,
David B. Grayden$^{1,\ast}$,
Visakan Kadirkamanathan$^{4}$
\\
\bf{1} Department of Electrical and Electronic Engineering, University of Melbourne, Melbourne, VIC, Australia
\\
\bf{2} The Bionic Ear Institute, East Melbourne, VIC, Australia
\\
\bf{3} Institute for Adaptive and Neural Computation, University of Edinburgh, Edinburgh, UK
\\
\bf{4} Department of Automatic Control and Systems Engineering, University of Sheffield, Sheffield, UK
\\
\bf{5} Department of Applied Physics and Applied Mathematics, Columbia University, New York, NY, USA
\\
\bf{6} Department of Systems and Control Engineering, University of Malta, Msida, MSD, Malta
\\
$\ast$ E-mail: dfreestone@bionicear.org
\end{flushleft}

@@

    	\caption{{\bf \textbf{(A)} Example of the neural field}. (\textbf{B}) Example of data generated by the full neural field model from the first 5 sensors. (\textbf{C}) Power spectral density of the data generated by an observation showing the typical $1/f^2$ characteristic of intracranial EEG.} 

@@

Using the model selection techniques described in Section~\ref{SpectralAnalysisSection}, the spatial frequency analysis was used to confirm that the spacing and width of the sensors was adequate to capture the significant dynamics of the field. Figure~\ref{fig:FieldFFT} shows the spatial frequency of the neural field. The cut-off frequency of the neural field is $\nu_c \approx 0.24$~Hz ($-3$~dB point). From equation~\ref{eq:MinimumSensorDistance}, this yielded a minimum sensor separation of $2.05$~mm. This confirms that the separation of $\Delta_{y} = 1.5$~mm was sufficient to prevent problems associated with spatial aliasing.

@@

The spatial frequency of the observed field is shown in Figure~\ref{fig:ObservationFreqAnal}. The cut-off frequency is $\nu_{cy} \approx 0.2$~Hz ($-3$~dB point). The observation cut-off frequency is lower than the field cut-off due to the low-pass filtering effect of the sensors. The spatial frequency response of the sensors can be seen in Figure~\ref{fig:SensorFunctionFreqResponse}. Using an oversampling parameter of $\rho_{\phi}=1$, the minimum distance between the field basis functions was calculated to be $2.5$~mm to satisfy Shannon's sampling criterion (from equation~\ref{eq:BasisFunctionSeparation}). The distance between basis functions was set to $\Delta_{\phi}=2.5$~mm, which is sufficient to capture the significant dynamics according to our criteria.
Given the basis function separation, the basis function width parameter was set to $\sigma_{\phi}^2=2.5$~mm. This allowed a smooth representation of the field by the basis function decomposition up to a spatial frequency $\mathbf{\nu_{c\phi}} \approx 0.12$~Hz, as described by equation~\ref{eq:CutoffFromBasisFuncWidth}. The band-limiting effect is due to the spatial frequency response of the basis functions as seen in Figure~\ref{fig:BasisFunctionFreqResponse}.

@@

		\label{fig:ObservationFreqAnal}\includegraphics{./Graph/SensorBasisObsFrequencyResponse.pdf}	

@@

@@

\subsection*{State and Parameter Estimates} 
\label{sec:state_and_param_results}
To demonstrate the performance of the state and parameter estimation algorithm, a Monte Carlo approach was employed where $100$ realisations of the data were generated. Each realisation consisted of $500$~ms of data (sampled at $1$~kHz) and the estimation was applied to the final $400$~ms, allowing the model's initial transients to die out. The initial state and parameters were unknown to the estimator. Eleven iterations of the estimation algorithm were used to ensure the parameters  converged. Figure~\ref{fig:ParametersConvergence} shows the rate of convergence, confirming the parameters converged, with the changes in parameters falling to less than $10^{-4}$ after six iterations.
\begin{figure}
        \centering
\includegraphics{./Graph/convergence.pdf}
\caption{{\bf The mean of the absolute error between true and estimated parameters averaged over 100 runs}. All parameters converge after six iterations, where the mean difference between errors in consecutive iterations falls below $10^{-4}$.}
\label{fig:ParametersConvergence}
\end{figure}

>>>>

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{A Data Driven Framework for Patient-Specific Neural Field Modelling}
}
% Insert Author names, affiliations and corresponding author email.
\\
Dean R. Freestone$^{1,2,3,\ast}$, 
Parham Aram$^{4}$, 
Michael Dewar$^{5}$,
Kenneth Scerri$^{6}$,
David B. Grayden$^{1}$,
Visakan Kadirkamanathan$^{4}$
\\
\bf{1} Department of Electrical and Electronic Engineering, University of Melbourne, Melbourne, VIC, Australia
\\
\bf{2} The Bionic Ear Institute, East Melbourne, VIC, Australia
\\
\bf{3} Institute for Adaptive and Neural Computation, University of Edinburgh, Edinburgh, UK
\\
\bf{4} Department of Automatic Control and Systems Engineering, University of Sheffield, Sheffield, UK
\\
\bf{5} Department of Applied Physics and Applied Mathematics, Columbia University, New York, NY, USA
\\
\bf{6} Department of Systems and Control Engineering, University of Malta, Msida, MSD, Malta
\\
$\ast$ E-mail: dfreestone@bionicear.org
\end{flushleft}

@@

    	\caption{{\bf Example of the spatio-temporal properties of the model and synthetically generated data.} \textbf{(A)} Example of the neural field at a single time instant. (\textbf{B}) Example of data generated by the full neural field model from the first 5 sensors. (\textbf{C}) Power spectral density of the data generated by an observation showing the typical $1/f^2$ characteristic of intracranial EEG.} 

@@

Using the model selection techniques described in Section~\ref{SpectralAnalysisSection}, the spatial frequency analysis was used to confirm that the spacing and width of the sensors was adequate to capture the significant dynamics of the field. Figure~\ref{fig:FFTTrueEstimate}(a) shows the spatial frequency of the neural field. The cut-off frequency of the neural field is $\nu_c \approx 0.24$~Hz ($-3$~dB point). From equation~\ref{eq:MinimumSensorDistance}, this yielded a minimum sensor separation of $2.05$~mm. This confirms that the separation of $\Delta_{y} = 1.5$~mm was sufficient to prevent problems associated with spatial aliasing.

@@

The spatial frequency of the observed field is shown in Figure~\ref{fig:ObservationFreqAnal}(c). The cut-off frequency is $\nu_{cy} \approx 0.2$~Hz ($-3$~dB point). The observation cut-off frequency is lower than the field cut-off due to the low-pass filtering effect of the sensors. The spatial frequency response of the sensors can be seen in Figure~\ref{fig:ObservationFreqAnal}(a). Using an oversampling parameter of $\rho_{\phi}=1$, the minimum distance between the field basis functions was calculated to be $2.5$~mm to satisfy Shannon's sampling criterion (from equation~\ref{eq:BasisFunctionSeparation}). The distance between basis functions was set to $\Delta_{\phi}=2.5$~mm, which is sufficient to capture the significant dynamics according to our criteria.
Given the basis function separation, the basis function width parameter was set to $\sigma_{\phi}^2=2.5$~mm. This allowed a smooth representation of the field by the basis function decomposition up to a spatial frequency $\mathbf{\nu_{c\phi}} \approx 0.12$~Hz, as described by equation~\ref{eq:CutoffFromBasisFuncWidth}. The band-limiting effect is due to the spatial frequency response of the basis functions as seen in Figure~\ref{fig:ObservationFreqAnal}(b). The spatial frequency of the reconstructed field from the basis function decomposition is shown in Figure~\ref{fig:FFTTrueEstimate}(b). This figure illustrates the band-limiting effect of the basis function decomposition.
\begin{figure}[!ht] 

@@

		\includegraphics{./Graph/SensorBasisObsFrequencyResponse.pdf}	

@@

	\label{fig:ObservationFreqAnal}

@@

%\begin{figure}[!ht]
%\begin{center}
%%\includegraphics[width=4in]{figure_name.2.eps}
%\end{center}
%\caption{
%{\bf Bold the first sentence.}  Rest of figure 2  caption.  Caption 
%should be left justified, as specified by the options to the caption 
%package.
%}
%\label{Figure_label}
%\end{figure}

\subsection*{State and Parameter Estimates} 
\label{sec:state_and_param_results}
To demonstrate the performance of the state and parameter estimation algorithm, a Monte Carlo approach was employed where $100$ realisations of the data were generated. Each realisation consisted of $500$~ms of data (sampled at $1$~kHz) and the estimation was applied to the final $400$~ms, allowing the model's initial transients to die out. The initial state and parameters were unknown to the estimator. Eleven iterations of the estimation algorithm were used to ensure the parameters  converged. Figure~\ref{fig:ParametersConvergence} shows the rate of convergence, confirming the parameters converged, with the changes in parameters falling to less than $10^{-4}$ after six iterations.
\begin{figure}
        \centering
\includegraphics{./Graph/convergence.pdf}
\caption{{\bf Convergence of parameters.} The mean of the absolute error between true and estimated parameters averaged over 100 runs. All parameters converge after six iterations, where the mean difference between errors in consecutive iterations falls below $10^{-4}$.}
\label{fig:ParametersConvergence}
\end{figure}
